---
title: "Ecological network control"
subtitle: "Supplementary Information"
author: "Fernando Cagua, Kate L. Wootton, Daniel B. Stouffer"
classoption: a4paper
csl: journal-of-ecology.csl
documentclass: artikel1
output: 
  bookdown::pdf_document2: 
    keep_tex: yes
    toc: false
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
bibliography: references.bib
---

\renewcommand\thefigure{S\arabic{figure}}    
\setcounter{figure}{0}   

\renewcommand\thetable{S\arabic{table}}    
\setcounter{table}{0}   

\renewcommand{\thesection}{S\arabic{section}}
\setcounter{section}{0}

\onehalfspacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE)
options(xtable.comment = FALSE, 
        xtable.caption.placement = 'bottom',
        xtable.size = 'small',
        booktabs = TRUE)
```

# Characteristics of the epirical networks {#empirical-networks}

```{r}
drake::loadd(metadata)
drake::loadd(network_properties)

net_prop <- metadata %>%
	dplyr::mutate(n_s_t = n_pla + n_pol) %>%
	dplyr::select(-n_pla, -n_pol) %>%
	dplyr::inner_join(network_properties, by = "net_name")

```

The networks studied had species richness ranging between `r min(net_prop$n_s_t)` and `r max(net_prop$n_s_t)` when considering only the largest component in each network). 
As shown by the network asymmetry *AS* [@bluthgen_specialization_2007], the networks had a low ratio of plants to pollinators overall. 
Furthermore, the networks had relatively low levels of nestedness [when measured using the quantitative version of the NODF index; @almeida-neto_straightforward_2011]. Details for each network can be found in the Table \@ref(tab:table-network-properties).

```{r table-network-properties, results = "asis", sanitize = FALSE}

net_prop %>%
	dplyr::select(-net_name, -count, -net) %>%
	dplyr::filter(study != "ballantyne") %>%
	dplyr::arrange(study, site, inv) %>% 
	dplyr::mutate(invader = dplyr::case_when(
				 	invader == "car" ~ "\\textit{Carpobrotus affine}",
				 	invader == "op" ~ "\\textit{Opuntia stricta}",
				 	invader == "imp" ~ "\\textit{Impatients grandulifera}",
				 	TRUE ~ "—"
				 ),
				 invader = dplyr::if_else(inv, invader, "—")) %>%
	dplyr::mutate(site = rep(1:n_distinct(site), each = 2),
				 inv = dplyr::if_else(inv, "yes", "no"),
				 study = dplyr::case_when(
				 	study == "bartomeus" ~ "Cap de Creus, Spain",
				 	TRUE ~ "Bristol, United Kingdom"
				 )) %>% 
  dplyr::select(site, invader, n_s_t, n_sp, n_pla, n_pol, connectance, web.asymmetry, weighted.NODF, study) %>%
  dplyr::rename_("$R$" = "n_s_t",
					"$n_p$" = "n_pla",
					"$n_a$" = "n_pol",
					"location" = "study",
					"$n_s$" = "n_sp",
					"$c$"= "connectance", 
					"NODF" = "weighted.NODF",
					# "$\\bar{a}$" = "interaction.strength.asymmetry",
					"$AS$" = "web.asymmetry") %>%
	xtable::xtable(caption = "Properties of the analysed plant-pollinator communities. Here show the number of species ($n_s$), the number of plants ($n_p$), the number of pollinators ($n_a$), the network connectance ($c$), the network asymmetry ($AS$), and the network nestedness (NODF index). All properties correspond to the network's largest component. British networks were assembled by Lopezaraiza-Mikel et al. (2007), Spanish were networks assembled by Bartomeus et al. (2008).", 
         label= "tab:table-network-properties") %>%
	print(#floating.environment = "sidewaystable", 
		comment = FALSE, sanitize.text.function = function(x){x}, include.rownames = FALSE, caption.placement = "top", booktabs = T)

# knitr::kable(meta)
```

# References
